Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation

### 2020-10-21 summary.Rmd
more simulations

* Number of vertices $n = 256, 512, 1024, 2048, 4096$
* Number of underlying communities $K = 2, 3, 4$
* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$, (i.e., each 
community has an equal probability of being drawn)
* Community labels 
$z_k \stackrel{iid}{\sim} Multinomial(\alpha_K, ..., \alpha_K)$
* Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
* Between-group popularities 
$\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$

@@

```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  geom_line(aes(x = n, y = med.err.mm,
                colour = 'Modularity maximization')) +
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
                    colour = 'Modularity maximization')) +
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```

@@

```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096. Simulations were repeated 40 times for each sample size. The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'misclustering count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC')) + 
  geom_line(aes(x = n, y = med.err.mm * n,
                colour = 'Modularity maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
                    colour = 'Modularity maximization')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() +
  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

@@

rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 4')) + 
  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                   colour = 'MLE-based')) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', colour = NULL) + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

@@

```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and subspace clustering (green) for sample sizes from 128 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  # scale_x_log10() + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2')) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC')) + 
  geom_line(aes(x = n, y = med.err.mm, colour = 'modularity maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm, 
                    colour = 'modularity maximization')) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  scale_y_log10()
```

@@

```{r}
dplyr::tibble(network = c('British MPs', 'Political blogs', 'DBLP'),
              vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
              `error (mod. max.)` = c(round(1 / 329, 3),
                                      round(61 / 1222, 3),
                                      round(62 / 2203, 3)),
              `error (SSC)` = round(1 - c(bmp.acc.ssc,
                                          pb.acc.ssc,
                                          dblp.acc.ssc), 3),
              `error (alg. 2)` = round(1 - c(bmp.acc, 
                                                  pb.acc,
                                                  dblp.acc), 3)) %>% 
  knitr::kable(caption = paste('Community detection error rates', 
                               'for modularity maximization,',
                               'sparse subspace clustering, ',
                               'and Algorithm 2.'))
```

@@

knitr::kable(data.frame(method = c('Algorithm 2', 'SSC', 
                                   'Modularity maximization'),
                        error = round(1 - c(hh.acc, hh.acc.ssc, hh.acc.mm), 3)),
             caption = paste('Community detection error rates for identifying',
                             'household religion.'))
```

>>>>

* Number of vertices $n = 256, 512, 1024, 2048, 4096$
* Number of underlying communities $K = 2, 3, 4$
* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$, (i.e., each 
community label has an equal probability of being drawn)
* Community labels 
$z_k \stackrel{iid}{\sim} Multinomial(\alpha_1, ..., \alpha_K)$
* Within-group popularities $\lambda^{(kk)} \stackrel{iid}{\sim} Beta(2, 1)$
* Between-group popularities 
$\lambda^{(kl)} \stackrel{iid}{\sim} Beta(1, 2)$ for $k \neq l$

@@

```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and sparse subspace clustering (green) for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm,
                colour = 'Modularity maximization')) +
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
                    colour = 'Modularity maximization'), width = .1) +
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') +
  scale_y_log10()
```

@@

```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'misclustering count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Algorithm 2'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm * n,
                colour = 'Modularity maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
                    colour = 'Modularity maximization'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10()
```

@@

rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 4'), width = .1) + 
  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                   colour = 'MLE-based'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', colour = NULL) + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

@@

```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (red) compared against modularity maximization (blue) and subspace clustering (green) for sample sizes from 128 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
clustering.df %>%
  dplyr::group_by(K, n) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'community detection error', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'Algorithm 2'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm, colour = 'Modularity maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm, 
                    colour = 'Modularity maximization'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') +
  scale_y_log10()
```

```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (red) compared against subspace clustering (blue) for sample sizes from 128 to 4096. Simulations were repeated 50 times for each sample size. The $y$-axis is shown on a logarithmic scale.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'misclustering count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Algorithm 2')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Algorithm 2'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm * n,
                colour = 'Modularity maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
                    colour = 'Modularity maximization'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10()
```

```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Log-log plot of the median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size.', fig.width = 10}
rmse.df <- readr::read_csv('rmse-k-imbalanced.csv')

rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 4'), width = .1) + 
  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                   colour = 'MLE-based'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', colour = NULL) + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

@@

```{r}
dplyr::tibble(Network = c('British MPs', 'Political blogs', 'DBLP'),
              Vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
              `Error (Mod. Max.)` = c(round(1 / 329, 3),
                                      round(61 / 1222, 3),
                                      round(62 / 2203, 3)),
              `Error (SSC)` = round(1 - c(bmp.acc.ssc,
                                          pb.acc.ssc,
                                          dblp.acc.ssc), 3),
              `Error (Alg. 2)` = round(1 - c(bmp.acc, 
                                                  pb.acc,
                                                  dblp.acc), 3)) %>% 
  knitr::kable(caption = paste('Community detection error rates', 
                               'for modularity maximization,',
                               'sparse subspace clustering, ',
                               'and Alg. 2.'))
```

@@

knitr::kable(data.frame(Method = c('Alg. 2', 'SSC', 
                                   'Mod. Max.'),
                        Error = round(1 - c(hh.acc, hh.acc.ssc, hh.acc.mm), 3)),
             caption = paste('Community detection error rates for identifying',
                             'household religion.'))
```
